---
title: "SV analysis"
output: null_document
---


# Setup
```{r}
library(pannagram)
library(ggplot2)

library(igraph)
library(GGally)
library(network)

library("cowplot")

Sys.setenv(PATH = paste("/Users/annaigolkina/micromamba/envs/pannagram/bin", 
                          Sys.getenv("PATH"), sep = ":"))
```

## Plot variables
```{r}
seed.value = 239
plot.size = 4
```

## Other variables
```{r}
min.len = 200

cov.cutoff = 85
cov.cutoff.strict = 85

min.copy = 2
```

## Paths
```{r}
path.data = '../examples/'
path.figures = '../images/families/'
```

## Files
```{r}

file.nestedness <- file.path(path.data, 'seq_sv_large_85_85.txt')
file.seqs = file.path(path.data, 'seq_sv_large.fasta')

```

# Read the data
```{r}
nestedness = read.table(file.nestedness, header = T)
seqs = readFasta(file.seqs)
```


# Filter by length and visualisation
```{r}

nestedness = filterNestedness(nestedness,
                               min.len = min.len)

edges.init = getGraphFromNestedness(nestedness, cov.cutoff = cov.cutoff)

g <- network(edges.init, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

set.seed(seed.value)
p <- ggnet2(g, label = F, edge.color = "black",
            node.size = 0.5,
            color = '#468B97',
            arrow.gap = 0.01, arrow.size = 2,
)
# p

# Save
i.plot = 1
name.output = sprintf("graph_%02d_init", i.plot)
savePNG(p, path = path.figures, name = name.output,
        width = plot.size, height = plot.size)
```


# Remain those nodes, that have at least two sequences + Length cutoff + Remain dominant edges
```{r}

nestedness.major = filterNestedness(nestedness,
                                     cov.cutoff = cov.cutoff,
                                     min.copy = min.copy)

edges.major = getGraphFromNestedness(nestedness.major)

g <- network(edges.major, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

set.seed(seed.value)
p <- ggnet2(g, label = F, edge.color = "black",
            node.size = 0.5,
            color = '#468B97',
            arrow.gap = 0.01, arrow.size = 2,
)
# p

# Save
i.plot = 2
name.output = sprintf("graph_%02d_no_singletons", i.plot)
savePNG(p, path = path.figures, name = name.output,
        width = plot.size, height = plot.size)
```

## Create a compact graph
```{r}

# --- Partition ---
components.info = getGraphComponents(edges.major) 
components = components.info$membership

# --- Compact ---
graph.compact = getGraphCompact(edges.major)
edges.compact = graph.compact$edges

# --- Common Components ---
svs.in.compact = graph.compact$nodes$name[graph.compact$nodes$node %in% edges.compact]
components.in.compact = data.frame(comp = components[svs.in.compact],
                              node = graph.compact$nodes[svs.in.compact,]$node)
components.in.compact = unique(components.in.compact)
if(nrow(components.in.compact) !=  length(unique(components.in.compact$node))) 
  stop('Wring assignment nodes to components')
components.in.compact = setNames(components.in.compact$comp,
                                 components.in.compact$node)


# --- Plot Compact ---
g <- network(edges.compact, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Add labels
g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(components.in.compact)] = as.character(components.in.compact)
g.label[duplicated(g.label)] = ''
g %v% "labels" = g.label

# Add cluster color
g %v% "partition" <- as.character(components.in.compact[g.names])

# Add cluster size
g %v% "size" <- graph.compact$nodes.size[g.names]

# Plot
set.seed(seed.value)
p.compact <- ggnet2(g, 
                    label = 'labels', label.size = 3,
                    edge.color = "grey30",
            node.size = 'size',
            color = 'partition',
            # arrow.gap = 0.01, 
            arrow.size = 2,
) + theme(legend.position = "none") + scale_color_viridis_d(name = "partition")

# --- Plot Partition ---
g <- network(edges.major, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Add labels
components.reduced = components
components.reduced[!(components.reduced %in% components.in.compact)] = NA

g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(components.reduced)] = as.character(components.reduced)
g.label[duplicated(g.label)] = ''
g.label[is.na(g.label)] = ''
g %v% "labels" = g.label

# Add cluster color
g %v% "partition" <- as.character(components.reduced[g.names])

# Plot
set.seed(seed.value)
p <- ggnet2(g, label = 'labels', 
            label.size = 3,
            edge.color = "grey",
            node.size = 0.3,
            color = 'partition',
) + theme(legend.position = "none") + 
  scale_color_viridis_d(name = "partition", na.value = "grey50")


p2 = cowplot::plot_grid(p.compact, p, ncol = 2)

i.plot = 3
name.output = sprintf("graph_%02d_compact", i.plot)
savePNG(p2, path = path.figures, name = name.output,
        width = plot.size * 3, height = plot.size * 1.5)

```


# Remove shortcut edges
```{r}

edges.no.shortcut <- filterEdgesShortcut(edges.major)
    
graph.compact = getGraphCompact(edges.no.shortcut)
edges.compact = graph.compact$edges

g <- network(edges.compact, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# # Add cluster size
# g %v% "size" <- graph.compact$nodes.size[g.names]
# set.seed(seed.value)
# p <- ggnet2(g, label = F, edge.color = "black",
#             node.size = 'size',
#             color = '#468B97',
#             arrow.gap = 0.01, arrow.size = 2,
# ) + theme(legend.position = "none")
# p 
# 

# Add labels
g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(components.in.compact)] = as.character(components.in.compact)
g.label[duplicated(g.label)] = ''
g %v% "labels" = g.label

# Add cluster color
g %v% "partition" <- as.character(components.in.compact[g.names])

# Add cluster size
g %v% "size" <- graph.compact$nodes.size[g.names]

# Plot
set.seed(seed.value)
p <- ggnet2(g, 
                    label = 'labels', label.size = 3,
                    edge.color = "grey30",
            node.size = 'size',
            color = 'partition',
            arrow.gap = 0.01, arrow.size = 2,
) + theme(legend.position = "none") + scale_color_viridis_d(name = "partition")



i.plot = 4
name.output = sprintf("graph_%02d_compact_no_shortcut", i.plot)
savePNG(p, path = path.figures, name = name.output,
        width = plot.size, height = plot.size)
```

# Remove dominants
```{r}

edges.no.dominant <- filterEdgesDominant(edges.no.shortcut)

```


# Plot compact
```{r}

# --- Partition ---
components.info = getGraphComponents(edges.no.dominant) 
components = components.info$membership

# --- Compact ---
graph.compact = getGraphCompact(edges.no.dominant)
edges.compact = graph.compact$edges

# --- Common Components ---
svs.in.compact = graph.compact$nodes$name[graph.compact$nodes$node %in% edges.compact]
components.in.compact = data.frame(comp = components[svs.in.compact],
                              node = graph.compact$nodes[svs.in.compact,]$node)
components.in.compact = unique(components.in.compact)
if(nrow(components.in.compact) !=  length(unique(components.in.compact$node))) 
  stop('Wring assignment nodes to components')
components.in.compact = setNames(components.in.compact$comp,
                                 components.in.compact$node)


# --- Plot Compact ---
g <- network(edges.compact, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Add labels
g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(components.in.compact)] = as.character(components.in.compact)
g.label[duplicated(g.label)] = ''
g %v% "labels" = g.label

# Add cluster color
g %v% "partition" <- as.character(components.in.compact[g.names])

# Add cluster size
g %v% "size" <- graph.compact$nodes.size[g.names]

# Plot
set.seed(seed.value)
p.compact <- ggnet2(g, 
                    label = 'labels', label.size = 3,
                    edge.color = "grey30",
            node.size = 'size',
            color = 'partition',
            # arrow.gap = 0.01, 
            arrow.size = 2,
) + theme(legend.position = "none") + scale_color_viridis_d(name = "partition")

# --- Plot Partition ---
g <- network(edges.no.dominant, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Add labels
components.reduced = components
components.reduced[!(components.reduced %in% components.in.compact)] = NA

g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(components.reduced)] = as.character(components.reduced)
g.label[duplicated(g.label)] = ''
g.label[is.na(g.label)] = ''
g %v% "labels" = g.label

# Add cluster color
g %v% "partition" <- as.character(components.reduced[g.names])

# Plot
set.seed(seed.value)
p <- ggnet2(g, label = 'labels', 
            label.size = 3,
            edge.color = "grey",
            node.size = 0.3,
            color = 'partition',
) + theme(legend.position = "none") + 
  scale_color_viridis_d(name = "partition", na.value = "grey50")


p2 = cowplot::plot_grid(p.compact, p, ncol = 2)

i.plot = 4
name.output = sprintf("graph_%02d_no_dominant", i.plot)
savePNG(p2, path = path.figures, name = name.output,
        width = plot.size * 3, height = plot.size * 1.5)

```


# Remove forks
```{r}

edges.major.no.forks <- solveForkNodes(edges = edges.no.dominant,
                                       seqs = seqs)

graph.compact = getGraphCompact(edges.major.no.forks)
edges.compact = graph.compact$edges

g <- network(edges.compact, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Add cluster size
g %v% "size" <- graph.compact$nodes.size[g.names]
set.seed(seed.value)
p <- ggnet2(g, label = F, edge.color = "black",
            node.size = 'size',
            color = '#468B97',
            arrow.gap = 0.01, arrow.size = 2,
) + theme(legend.position = "none")
p 

i.plot = 5
name.output = sprintf("graph_%02d_compact_no_forks", i.plot)
savePNG(p, path = path.figures, name = name.output,
        width = plot.size, height = plot.size)

```

# Solve Umbrelas
```{r}

edges.major.no.umbrella <- solveUmbrellaNodes(edges = edges.major.no.forks,
                                         seqs = seqs, show.echo=T)


# --- Partition ---
components.info = getGraphComponents(edges.major.no.umbrella) 
components = components.info$membership

# --- Compact ---
graph.compact = getGraphCompact(edges.major.no.umbrella)
edges.compact = graph.compact$edges

# --- Common Components ---
svs.in.compact = graph.compact$nodes$name[graph.compact$nodes$node %in% edges.compact]
components.in.compact = data.frame(comp = components[svs.in.compact],
                              node = graph.compact$nodes[svs.in.compact,]$node)
components.in.compact = unique(components.in.compact)
if(nrow(components.in.compact) !=  length(unique(components.in.compact$node))) 
  stop('Wring assignment nodes to components')
components.in.compact = setNames(components.in.compact$comp,
                                 components.in.compact$node)


# --- Plot Compact ---
g <- network(edges.compact, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Add labels
g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(components.in.compact)] = as.character(components.in.compact)
g.label[duplicated(g.label)] = ''
g %v% "labels" = g.label

# Add cluster color
g %v% "partition" <- as.character(components.in.compact[g.names])

# Add cluster size
g %v% "size" <- graph.compact$nodes.size[g.names]

# Plot
set.seed(seed.value)
p.compact <- ggnet2(g, 
                    label = 'labels', label.size = 3,
                    edge.color = "black",
            node.size = 'size',
            color = 'partition',
            arrow.gap = 0.01, arrow.size = 2,
) + theme(legend.position = "none") + scale_color_viridis_d(name = "partition")
p.compact

# --- Plot Partition ---
g <- network(edges.major.no.umbrella, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Add labels
components.reduced = components
components.reduced[!(components.reduced %in% components.in.compact)] = NA

g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(components.reduced)] = as.character(components.reduced)
g.label[duplicated(g.label)] = ''
g.label[is.na(g.label)] = ''
g %v% "labels" = g.label

# Add cluster color
g %v% "partition" <- as.character(components.reduced[g.names])

# Plot
set.seed(seed.value)
p <- ggnet2(g, label = 'labels', 
            label.size = 3,
            edge.color = "grey",
            node.size = 0.5,
            color = 'partition',
) + theme(legend.position = "none") + 
  scale_color_viridis_d(name = "partition", na.value = "grey50")
p



p2 = cowplot::plot_grid(p.compact, p, ncol = 2)

i.plot = 6
name.output = sprintf("graph_%02d_no_umbrella", i.plot)
savePNG(p2, path = path.figures, name = name.output,
        width = plot.size * 4, height = plot.size * 2)

```

# Components
```{r}

components.info = getGraphComponents(edges.major.no.umbrella) 
components = components.info$membership

g <- network(edges.major.no.umbrella, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Add cluster color
g %v% "partition" <- as.character(components[g.names])

# labels
g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(components)] = as.character(components)
g.label[duplicated(g.label)] = ''
g %v% "labels" = g.label

set.seed(seed.value)
p <- ggnet2(
  g,
  label = 'labels',
  label.size = 3,
  edge.color = "grey70",
  node.size = 0.1,
  color = "partition",
  arrow.gap = 0.01,
  arrow.size = 2
) + theme(legend.position = "none") +
  scale_color_viridis_d(name = "partition")

i.plot = 7
name.output = sprintf("graph_%02d_partition", i.plot)
savePNG(p, path = path.figures, name = name.output,
        width = plot.size * 2, height = plot.size * 2)

# rm(components.info)
# rm(components)
```

# Stop
```{r}
stop()
```


# Put sequences back
```{r}
dominant.effect = 0.7 

edges.solved = putEdgesBack(edges = edges.major.no.umbrella, 
                            edges.init = edges.init,
                            dominant.effect = dominant.effect, 
                            show.echo=T)
nrow(edges.solved)
edges.solved <- filterEdges(edges.solved, min.comp.size = 3)
nrow(edges.solved)
edges.solved <- filterEdgesShortcut(edges.solved, show.echo = T)
nrow(edges.solved)

```

## Remove umbrella & Forks
```{r}

edges.solved.filter = edges.solved

nrow(edges.solved.filter)
edges.solved.filter <- solveUmbrellaNodes(edges = edges.solved.filter,
                                         seqs = seqs, show.echo=T)
nrow(edges.solved.filter)
edges.solved.filter <- solveForkNodes(edges = edges.solved.filter,
                                         seqs = seqs, show.echo = T)

nrow(edges.solved.filter)
edges.solved.filter <- filterEdgesDominant(edges.solved.filter)
nrow(edges.solved.filter)

```


```{r}
components.info = getGraphComponents(edges.solved.filter) 
components = components.info$membership

g <- network(edges.solved.filter, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Add cluster color
g %v% "partition" <- as.character(components[g.names])

# labels
g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(components)] = as.character(components)
g.label[duplicated(g.label)] = ''

g %v% "labels" = g.label

set.seed(seed.value)
p <- ggnet2(
  g,
  label = 'labels',
  label.size = 3,
  edge.color = "grey70",
  node.size = 0.1,
  color = "partition",
  arrow.gap = 0.005,
  arrow.size = 2
) + theme(legend.position = "none") +
  scale_color_viridis_d(name = "partition")

i.plot = 8
name.output = sprintf("graph_%02d_solved", i.plot)
savePNG(p, path = path.figures, name = name.output,
        width = plot.size * 2, height = plot.size * 2)


# rm(components.info)
# rm(components)

```

## Compact
```{r}

graph.compact = getGraphCompact(edges.solved.filter)
edges.compact = graph.compact$edges 

graph.compact$nodes$comp = components[graph.compact$nodes$name]

components.nodes = unique(graph.compact$nodes[, c('node', 'comp')])
rownames(components.nodes) = NULL
components.nodes = setNames(components.nodes$comp, components.nodes$node)

g <- network(edges.compact, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# labels
g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(components.nodes)] = as.character(components.nodes)
g.label[duplicated(g.label)] = ''

g %v% "labels" = g.label

# Add cluster size
g %v% "size" <- graph.compact$nodes.size[g.names]
set.seed(seed.value)
p.compact <- ggnet2(g, label = 'labels', 
                    label.size = 3,
                    edge.color = "black",
            node.size = 'size',
            color = '#468B97',
            arrow.gap = 0, arrow.size = 2,
) + theme(legend.position = "none")
p.compact


i.plot = 9
name.output = sprintf("graph_%02d_solved_compact", i.plot)
savePNG(p.compact, path = path.figures, name = name.output,
        width = plot.size * 2, height = plot.size * 2)

```


# Stop
```{r}
stop()
```


# Separate components!
```{r}


components.info = getGraphComponents(edges.solved.filter) 
components = components.info$membership

i.comp = 40

seqs.names.comp = names(components)[components == i.comp]


seqs.comp <- getComponentSequences(seqs.names.comp = seqs.names.comp,
                              seqs, 
                              nestedness)

# dotplot(seqs.comp[1], seqs.comp[2])


hist(nchar(seqs.comp), 50)


```

## plot Families
```{r}

edges.fam = getFmilyEdges(edges.solved.filter, components, i.comp = i.comp)

g <- network(edges.fam, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

set.seed(seed.value)
p <- ggnet2(g, label = F, edge.color = "black",
            node.size = 1,
            color = '#468B97',
            arrow.gap = 0.01, arrow.size = 2,
)
p


# # --- Compact ---
# graph.compact = getGraphCompact(edges.fam)
# edges.compact = graph.compact$edges
# 
# # --- Plot Compact ---
# g <- network(edges.compact, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
# g.names = network.vertex.names(g)
# 
# # Add labels
# g %v% "labels" = g.names
# 
# # Add cluster size
# g %v% "size" <- graph.compact$nodes.size[g.names]
# 
# # Plot
# set.seed(seed.value)
# p.compact <- ggnet2(g, 
#                     label = 'labels', label.size = 3,
#                     edge.color = "grey30",
#             node.size = 'size',
#             arrow.gap = 0.1,
#             arrow.size = 2,
# ) + theme(legend.position = "none") 
# p.compact

# 
# sv1 = graph.compact$nodes$name[graph.compact$nodes$node == 'N15']
# 
# g <- network(edges.fam, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
# g.names = network.vertex.names(g)
# g %v% "highlight" <- g.names %in% sv1
# p <- ggnet2(
#   g,
#   label = FALSE,
#   edge.color = "black",
#   node.size = 1,
#   color = "highlight",
#   arrow.gap = 0.01,
#   arrow.size = 2
# ) +
#   scale_color_manual(
#     values = c(`TRUE` = "red", `FALSE` = "#468B97"),
#     guide = "none"
#   )
# 
# p
# 
# 
# graph.compact$nodes$node[graph.compact$nodes$name == 'SVgr_3_id_24499|361']
# 
# sv1 = graph.compact$nodes$name[graph.compact$nodes$node == 'N15']
# sv2 = graph.compact$nodes$name[graph.compact$nodes$node == 'N14']
# 
# edges.init[(edges.init[,1] %in% sv2) & (edges.init[,2] %in% sv1),]
# 
# dotplot(seqs[sv1[1]], seqs[sv2[1]])
# 
# dotplot(seqs["SVgr_1_id_043169|369"], seqs["SVgr_1_id_133369|3181"])

```


## Alignment
```{r}


path.mafft = "~/Sources/pannagram_test/test_ref_free_acc_old/features/sv/tmp/mafft/"

  tmp.fasta = paste(path.mafft, 'tmp_',i.comp,'.fasta', sep = '')
  aln.fasta = paste(path.mafft, 'aln_',i.comp,'.fasta', sep = '')
  
  writeFasta(seqs.comp, tmp.fasta)
  
  pokaz('Mafft...')
  system(paste('mafft --op 5 --quiet --maxiterate 100 ', tmp.fasta, '>', aln.fasta,  sep = ' '))
  
  pokaz('Plot...')
  alignment = readFasta(aln.fasta)
  
  msaplot(alignment)
  p = msadiff(alignment)
  
  p

  dotplot(alignment[20], alignment[30])
  
  ntplot(alignment[1],nt.separate = T)
```



```{r}
aln.reduces = reduceAlnFreq(alignment, show.echo = T)
msadiff(aln.reduces)

aln.reduces.flank = reduceAlnFlank(aln.reduces, show.echo = T)
msadiff(aln.reduces.flank)
msaplot(aln.reduces.flank)

ntplot(aln.reduces.flank[1])

s.cons = mx2cons(aln2mx(aln.reduces.flank))

dotplot(s.cons, s.cons)
```



```{r}
orf.info = orfFinder(seqs.comp[1])
orfplot(orf.info$pos)

orf.info$orf[1]


```





# Plot Families
```{r}

components.info = getGraphComponents(edges.solved.filter) 
components = components.info$membership

path.fam = "~/Sources/alignments/mosqioto/families/"
for(i.comp in 1:max(components)){
  
  pokaz(i.comp)

  seqs.names.comp = names(components)[components == i.comp]
  
  seqs.comp <- getComponentSequences(seqs.names.comp = seqs.names.comp,
                                seqs, 
                                nestedness)

  path.mafft = "~/Sources/alignments/mosqioto/families/tmp/"

  tmp.fasta = paste(path.mafft, 'tmp_',i.comp,'.fasta', sep = '')
  aln.fasta = paste(path.mafft, 'aln_',i.comp,'.fasta', sep = '')
  
  if(!(file.exists("aln.fasta") && file.info("aln.fasta")$size > 0) ) {
  
    
    writeFasta(seqs.comp, tmp.fasta)
    
    pokaz('Mafft...')
    system(paste('mafft --op 5 --quiet --maxiterate 100 ', tmp.fasta, '>', aln.fasta,  sep = ' '))
    
  }
  pokaz('Plot...')
  alignment = readFasta(aln.fasta)
  
  msaplot(alignment)
  p = msadiff(alignment)
  
  savePNG(p, path = path.fam, name = paste0('fam_', i.comp), width = 6, height = 5)
}

```

