---
title: "SV analysis"
output: md_document
---

# Setup
## Functions
```{r}
library(ggplot2)
library(cowplot)
library(rhdf5)
library(pannagram)
library(igraph)
library(ggnet)
library(network)

Sys.setenv(PATH = paste("/Users/annaigolkina/micromamba/envs/pannagram/bin", 
                          Sys.getenv("PATH"), sep = ":"))

if (!require("igraph")) install.packages("igraph")
if (!require("GGally")) install.packages("GGally")
if (!require("network")) install.packages("network")

library(igraph)
library(GGally)
library(network)
library(gridExtra)


```

## Paths
```{r}
s.species = 'b_juncea'
path.sv = '/Users/annaigolkina/Library/CloudStorage/OneDrive-Personal/iglab/projects/ufo_proj/07_sv/b_juncea/'

s.species = 'b_rapa'
path.sv = '/Users/annaigolkina/Library/CloudStorage/OneDrive-Personal/iglab/projects/ufo_proj/07_sv/b_rapa/'

s.species = 'camelina_sativa'
path.sv = '/Volumes/Samsung_T5/vienn/test/camelina_sativa/'

s.species = 'b_nigra'
path.sv = '/Users/annaigolkina/Library/CloudStorage/OneDrive-Personal/iglab/projects/ufo_proj/07_sv/b_nigra/'

s.species = 'cucumber'
path.sv = '/Users/annaigolkina/Library/CloudStorage/OneDrive-Personal/iglab/projects/ufo_proj/07_sv/cucumber/'

s.species = 'auqilegia'
path.sv = '/Users/annaigolkina/Library/CloudStorage/OneDrive-Personal/iglab/projects/ufo_proj/07_sv/auqilegia/'


s.species = 'rleg'
path.sv = '/Volumes/Samsung_T5/vienn/test/manuals/tmp/sv/'
```


```{r}
path.figures = paste(path.sv, 'figures/', sep = '')

path.nad = paste(path.sv, 'nad/', sep = '')

if (!dir.exists(path.figures)) dir.create(path.figures)
if (!dir.exists(path.nad)) dir.create(path.nad)

```

## Variables
```{r}
sim.cutoff = 0.85
sim.cutoff = 0.9
f.visualise = F
```

## Binning
```{r}

bins <- c(0, 100, 200, 400, 800, 1000, 1500, 3000, 5000, 7000, Inf)
labels <- c("0-100", "100-200", "200-400", "400-800", "800-1000", "1000-1500", "1500-3000", "3000-5000", "5000-7000", "7000+")


color.pal <- c("0-100" = "#1f77b4", "100-200" = "#50B498", "200-400" = "#2ca02c", 
            "400-800" = "#bcbd22", "800-1000" = "#ff7f0e", "1000-1500" = "#d62728", "1500-3000" = "#9467bd",
            "3000-5000" = "#e377c2", "5000-7000" = "#8c564b", "7000+" = "#7f7f7f")


```


# SV-graph
## Reading and constructing
```{r}

res.cover.file = 'seq_sv_big_on_sv_cover.rds'
res.nest = readRDS(paste(path.sv, res.cover.file, sep = ''))
g.content = getGraphFromBlast(res.nest = res.nest, sim.cutoff = sim.cutoff, collapse = T)

```
## Take only copies as proxy to the Mobilome
```{r}
min.cnt = 2
node.remain = g.content$nodes.traits$node[g.content$nodes.traits$cnt >=  min.cnt]
sv.remain = g.content$nodes$name[g.content$nodes$node %in% node.remain]

res.nest.remain = res.nest[(res.nest$V1 %in% sv.remain) & 
                             (res.nest$V8 %in% sv.remain),]

g.content = getGraphFromBlast(res.nest = res.nest.remain, sim.cutoff = sim.cutoff, collapse = F)

sum(g.content$nodes.traits$cnt > 100)

x = unique(c(g.content$edges))
length(x)

```


# Plot simple graph
```{r}
if(f.visualise){
  g <- network(g.content$edges, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
  b.graph.names = network.vertex.names(g)

  set.seed(239)
  p.refined <- ggnet2(g, label = F, edge.color = "black",
              # node.size = g.nodes.cnt[b.graph.names],
              node.size = 1,
              color = '#468B97',
              arrow.gap = 0.01, arrow.size = 2,
              # color = g.nodes.fam[b.graph.names],
              # palette = fam.palette,
              # mode = "kamadakawai"
              )
  
  p.refined
}

```

# Remove components
## Components
```{r}
g.comp <- getGraphComponents(g.content$edges)
memb.comp = g.comp$membership
g.content$edges.small = g.content$edges
```

## Remove one the biggest component and some the smallest
```{r}
# # Remove big
# id.big = which.max(g.comp$csize)
# sv.name.big = names(g.comp$membership)[g.comp$membership == id.big]
# g.content$edges.small = g.content$edges.small[!(g.content$edges[,1] %in% sv.name.big),]
```

## Remove some the smallest
```{r}
# Remove small
min.cl.size = 4
idx.moderate = which(g.comp$csize >= min.cl.size)
sv.name.moderate = names(g.comp$membership)[g.comp$membership %in% idx.moderate]
g.content$edges.small = g.content$edges.small[(g.content$edges.small[,1] %in% sv.name.moderate),]
```

## Collapse small graph
```{r}
names.tmp = unique(c(g.content$edges.small))
res.nest.small = res.nest[(res.nest$V1 %in% names.tmp) & (res.nest$V8 %in% names.tmp),]

sum(!(res.nest.small$V8 %in% names.tmp))

g.content.small = getGraphFromBlast(res.nest = res.nest.small, sim.cutoff = sim.cutoff, collapse = F)
```


### Fam Small
```{r}

v.sim = readRDS(paste0(path.sv, 'seq_sv_big_on_set_cover.rds'))

# TE family
v.sim$fam = sapply(v.sim$V8, function(s) strsplit(s, '\\|')[[1]][9])
v.sim$fam[v.sim$fam %in% c('DNA/Pogo', 'DNA/Tc1', 'DNA/Harbinger', 'DNA/En-Spm',
                                 'DNA', 'DNA/Mariner')] = 'DNA-TE'
v.sim$fam[v.sim$fam %in% c('RathE1_cons', 'RathE2_cons', 'RathE3_cons')] = 'RathE1/2/3_cons'
v.sim$fam[v.sim$fam %in% c('LINE/L1', 'LINE?')] = 'LINE'
v.sim$fam[v.sim$fam %in% c('RC/Helitron')] = 'Helitron'

family.stat = tapply(v.sim$fam, v.sim$V1, function(s){
    s0 = s
    s = unique(s)
    if(length(s) == 1) return(s)
    s = setdiff(s, 'TEG')
    if(length(s) == 1) return(s)
    s = setdiff(s, 'Unassigned')
    if(length(s) == 1) return(s)
    
    s0 = s0[s0 != 'TEG']
    s0 = s0[s0 != 'Unassigned']
    cnt0 = table(s0)
    if( (4 * max(cnt0)) >= (3 * length(s0))){
      # return(paste(names(cnt0)[cnt0 == max(cnt0)], 'm', sep = '_'))
      return(names(cnt0)[cnt0 == max(cnt0)])
    }
    
    if(length(s) == 2) {
      if(sum(s == 'Helitron') == 1){
        return(paste('Helitron', setdiff(s, 'Helitron'), sep = '/'))
      }
    }
    
    return('Mix')
  } )
  
family.stat[family.stat == 'Helitron/DNA/MuDR'] = 'DNA/MuDR' 

idx = g.content.small$nodes$name %in% names(family.stat)
g.content.small$nodes$fam = ''
g.content.small$nodes$fam[idx] = family.stat[g.content.small$nodes$name[idx]]


node.fam = tapply(g.content.small$nodes$fam, g.content.small$nodes$node, function(s) {
  s.init = s
  s = s[s != '']
  if(length(s) == 0){
    return('')
  } else if(length(s) != 1){
    cnt = table(s.init)
    s = names(cnt)[cnt == max(cnt)]
    if(length(s) == 1){
      return(s)
    } else {
      return('Mix')  
    }
  } else {
    return(s)
  }
})
node.fam = node.fam[node.fam != '']

idx = g.content.small$nodes.traits$node %in% names(node.fam)
g.content.small$nodes.traits$fam = ''
g.content.small$nodes.traits$fam[idx] = node.fam[g.content.small$nodes.traits$node[idx]]


```


### Fam Full
```{r}

g.content.small = g.content

v.sim = readRDS(paste0(path.sv, 'seq_sv_big_on_set_cover.rds'))

# TE family
v.sim$fam = sapply(v.sim$V8, function(s) strsplit(s, '\\|')[[1]][9])
v.sim$fam[v.sim$fam %in% c('DNA/Pogo', 'DNA/Tc1', 'DNA/Harbinger', 'DNA/En-Spm',
                                 'DNA', 'DNA/Mariner')] = 'DNA-TE'
v.sim$fam[v.sim$fam %in% c('RathE1_cons', 'RathE2_cons', 'RathE3_cons')] = 'RathE1/2/3_cons'
v.sim$fam[v.sim$fam %in% c('LINE/L1', 'LINE?')] = 'LINE'
v.sim$fam[v.sim$fam %in% c('RC/Helitron')] = 'Helitron'

family.stat = tapply(v.sim$fam, v.sim$V1, function(s){
    s0 = s
    s = unique(s)
    if(length(s) == 1) return(s)
    s = setdiff(s, 'TEG')
    if(length(s) == 1) return(s)
    s = setdiff(s, 'Unassigned')
    if(length(s) == 1) return(s)
    
    s0 = s0[s0 != 'TEG']
    s0 = s0[s0 != 'Unassigned']
    cnt0 = table(s0)
    if( (4 * max(cnt0)) >= (3 * length(s0))){
      # return(paste(names(cnt0)[cnt0 == max(cnt0)], 'm', sep = '_'))
      return(names(cnt0)[cnt0 == max(cnt0)])
    }
    
    if(length(s) == 2) {
      if(sum(s == 'Helitron') == 1){
        return(paste('Helitron', setdiff(s, 'Helitron'), sep = '/'))
      }
    }
    
    return('Mix')
  } )
  
family.stat[family.stat == 'Helitron/DNA/MuDR'] = 'DNA/MuDR' 

idx = g.content.small$nodes$name %in% names(family.stat)
g.content.small$nodes$fam = ''
g.content.small$nodes$fam[idx] = family.stat[g.content.small$nodes$name[idx]]


node.fam = tapply(g.content.small$nodes$fam, g.content.small$nodes$node, function(s) {
  s.init = s
  s = s[s != '']
  if(length(s) == 0){
    return('')
  } else if(length(s) != 1){
    cnt = table(s.init)
    s = names(cnt)[cnt == max(cnt)]
    if(length(s) == 1){
      return(s)
    } else {
      return('Mix')  
    }
  } else {
    return(s)
  }
})
node.fam = node.fam[node.fam != '']

idx = g.content.small$nodes.traits$node %in% names(node.fam)
g.content.small$nodes.traits$fam = ''
g.content.small$nodes.traits$fam[idx] = node.fam[g.content.small$nodes.traits$node[idx]]


```


### Save
```{r}
write.table(g.content.small$edges, paste0(path.nad, 'graph_collapse_',s.species,'_full.txt'), quote = F,
            row.names = F, col.names = F, sep = '\t')
write.table(g.content.small$nodes.traits, paste0(path.nad, 'graph_collapse_node_traits_',s.species,'_full.txt'), 
            quote = F, col.names = T, sep = '\t')
write.table(g.content.small$nodes, paste0(path.nad, 'graph_collapse_nodes_',s.species,'_full.txt'), quote = F, 
            row.names = F, col.names = T, sep = '\t')

```



# Plot small components
```{r}

g.sm <- network(g.content.small$edges, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
b.graph.names = network.vertex.names(g.sm)
set.seed(239)
p.sm <- ggnet2(g.sm, label = F, edge.color = "black",
            # node.size = g.nodes.cnt[b.graph.names],
            node.size = 1,
            color = '#468B97',
            arrow.gap = 0.01, arrow.size = 2,
            # color = g.nodes.fam[b.graph.names],
            # palette = fam.palette,
            # mode = "kamadakawai"
            )

p.sm

```

## Partition
```{r}

# Создание графа через igraph
edges <- g.content.small$edges  # Ваши данные о рёбрах
igraph_g <- graph_from_edgelist(as.matrix(edges), directed = TRUE)

igraph_g <- as.undirected(igraph_g, mode = "collapse")

# Применение метода Лувена
louvain_result <- cluster_louvain(igraph_g)

# Получение кластеров для каждого узла
partition <- membership(louvain_result)

# Конвертируем igraph в объект network
g <- network(edges, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Добавляем атрибут "partition" для узлов
g %v% "partition" <- as.character(partition[g.names])

# Визуализация с ggnet2
p.refined <- ggnet2(
  g,
  label = FALSE,
  edge.color = "black",
  node.size = 1,
  color = "partition",  # Используем найденные сообщества
  arrow.gap = 0.01,
  arrow.size = 2
) + theme(legend.position = "none")


p.refined + scale_color_viridis_d(name = "Partition")

```
## Remove nodes, that are connecting two different partitions
```{r}

# Рассчитать центральность посредничества
bridge_scores <- igraph::betweenness(igraph_g, directed = TRUE, normalized = TRUE)
hist(bridge_scores)
threshold <- 0.005
bridges <- which(bridge_scores > threshold)


g %v% "bridge" <- (g.names %in% names(bridges)) * 1 + 1

# Визуализация с ggnet2
p.refined <- ggnet2(
  g,
  label = FALSE,
  node.shape = 'bridge',
  # node.size = 'bridge',
  node.size = 1,
  color = "partition",  # Используем найденные сообщества
  arrow.gap = 0.01,
  arrow.size = 2
) + theme(legend.position = "none") + 
  scale_color_viridis_d(name = "Partition")

p.refined


```

# Separate components
```{r}

edges = g.content.small$edges

calcClusterFraction <- function(edges, node.b) {
  edges.b = edges[rowSums(edges == node.b) > 0,,drop=F]
  node.b.connect = setdiff(c(edges.b), node.b)
  node.b.connect.val = partition[node.b.connect]
  cnt = table(node.b.connect.val)
  p = max(cnt) / length(node.b.connect.val)
  return(p)
}

edges = g.content.small$edges
igraph_g <- graph_from_edgelist(as.matrix(edges), directed = TRUE)
igraph_g <- as.undirected(igraph_g, mode = "collapse")
louvain_result <- cluster_louvain(igraph_g)
partition <- membership(louvain_result)
bridge_scores <- igraph::betweenness(igraph_g, directed = TRUE, normalized = TRUE)
node.bridges = names(bridge_scores)[order(-bridge_scores)]

result <- sapply(node.bridges[1:round(length(node.bridges)/4)], 
                 function(v) calcClusterFraction(edges, v))

node.remove = names(result)[which(result < 0.95)]

edges = edges[!(edges[,1] %in% node.remove) & !(edges[,2] %in% node.remove),,drop=F]

```

## Plot
```{r}


igraph_g <- graph_from_edgelist(as.matrix(edges), directed = TRUE)

igraph_g <- as.undirected(igraph_g, mode = "collapse")

# Применение метода Лувена
louvain_result <- cluster_louvain(igraph_g)

# Получение кластеров для каждого узла
partition <- membership(louvain_result)

# Конвертируем igraph в объект network
g <- network(edges, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g)

# Добавляем атрибут "partition" для узлов
g %v% "partition" <- as.character(partition[g.names])

# Визуализация с ggnet2
p.refined <- ggnet2(
  g,
  label = FALSE,
  edge.color = "black",
  node.size = 1,
  color = "partition",  # Используем найденные сообщества
  arrow.gap = 0.01,
  arrow.size = 2
) + theme(legend.position = "none")


p.refined + scale_color_viridis_d(name = "Partition")


```



# Only the inresesting length range
```{r}

len.max = 400
len.min = 200

sv.len.tmp = as.numeric(sapply(names(g.comp$membership), function(s) strsplit(s, '\\|')[[1]][2]))
tmp = tapply(sv.len.tmp, g.comp$membership, function(a) sum((a <= len.min) | (a >= len.max)))
comp.remove = tmp[tmp != 0]

comp.get = setdiff(1:g.comp$no, as.numeric(names(comp.remove)))

sv.get = names(g.comp$membership)[g.comp$membership %in% comp.get]
g.content$edges.small = g.content$edges.small[(g.content$edges.small[,1] %in% sv.get),]

dim(g.content$edges.small)

# g.content$edges.small = g.content$edges

```

#### Plot small component
```{r}

g.sm <- network(g.content$edges.small, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
b.graph.names = network.vertex.names(g.sm)
set.seed(239)
p.sm <- ggnet2(g.sm, label = F, edge.color = "black",
            # node.size = g.nodes.cnt[b.graph.names],
            node.size = 1,
            color = '#468B97',
            arrow.gap = 0.01, arrow.size = 2,
            # color = g.nodes.fam[b.graph.names],
            # palette = fam.palette,
            # mode = "kamadakawai"
            )

p.sm

```

### With labels
```{r}

# edges = g.content$edges.small

g.sm <- network(edges, matrix.type = "edgelist", ignore.eval = FALSE, directed = TRUE)
g.names = network.vertex.names(g.sm)

g.comp <- getGraphComponents(edges)
memb.comp = g.comp$membership

g.label = rep('', length(g.names))
names(g.label) = g.names
g.label[names(memb.comp)] = as.character(memb.comp)
g.label[duplicated(g.label)] = ''

g.sm %v% "labels" = g.label

g.names.len = as.numeric(sapply(g.names, function(s) strsplit(s, '\\|')[[1]][2]))
g.names.len.bin <- as.character(cut(g.names.len, breaks = bins, right = FALSE, labels = labels))
g.sm %v% "colors" = as.character(g.names.len.bin)


p <- ggnet2(g.sm, label = 'labels', edge.color = "grey70", 
            # node.size = g.content$nodes.traits[g.part.sub.big.names,]$cnt, 
            # color = g.content$nodes.traits[g.part.sub.big.names,]$type,
            # mode = 'kamadakawai',
            node.size = 1,
            color = 'colors',
            palette = color.pal
            ) + 
  # theme(legend.position = "none") +
  guides(size = F)
p


png(paste0(path.figures, 'graph_small_', s.species,'.png', sep = ''), width = 6, height = 6, units = "in", res = 300)
print(p)    
dev.off()

```

```{r}
cluster_louvain(g.sm)
```


# Get sequences from every component
## Read sequence
```{r}
seqs = readFasta(paste0(path.sv, 'seq_sv_big.fasta'))
```

## Save all components
```{r}

comp.used = unique(g.comp$membership[b.graph.names])

s.ufo = c()

for(i.comp in comp.used){
  pokaz(i.comp)
  i.seq = which(memb.comp == i.comp)
  names.comp = names(memb.comp)[i.seq]
  s = seqs[names.comp[1]]

  s.ufo[paste0('comp', i.comp)] = s
}

names(s.ufo) = paste0(s.species, '|', names(s.ufo))

writeFasta(s.ufo, paste0(path.sv, 'ufo_',s.species,'.fasta'))


```

```{r}
stop()
```

# Alignment
```{r}


save(list = ls(), file = paste0(path.figures, "tmp_workspace_s.RData"))

path.mafft = '/Users/annaigolkina/Library/CloudStorage/OneDrive-Personal/iglab/projects/ufo_proj/07_sv/sv_cucur/tmp_mafft/'

for(i.comp in 1:max(memb.comp)){
  pokaz('Component', i.comp)
  i.seq = which(memb.comp == i.comp)
  
  names.comp = names(memb.comp)[i.seq]
  pokaz('Number of sequences', length(names.comp))

  sv.name.target = names.comp
  
  seqs.target = seqs[sv.name.target]
  seqs.target.len = as.numeric(sapply(names(seqs.target), function(s) strsplit(s, '\\|')[[1]][2]))
  seqs.target = seqs.target[order(-seqs.target.len)]
  sv.name.target = names(seqs.target)
  
  
  # Find orientations of sequences
  
  bl.target = res.nest[(res.nest$V1 %in% sv.name.target) & (res.nest$V8 %in% sv.name.target),]
  orientation.target = rep('.', length(seqs.target))
  names(orientation.target) = sv.name.target
  
  # Define the first orientation by the longest ORF
  s = seqs[sv.name.target[1]]
  orf.res = orfFinder(s)
  orientation.target[1] = orf.res$pos$strand[1]
  
  dir.seq = c('-', '+')
  
  for(i in 1:(length(orientation.target) - 1)){
    bl.tmp = bl.target[bl.target$V1 == sv.name.target[i],]
    bl.tmp = bl.tmp[!duplicated(bl.tmp$V8),]
    bl.tmp = bl.tmp[orientation.target[bl.tmp$V8] == '.',]
    if(nrow(bl.tmp) == 0) next
    
    for(irow in 1:nrow(bl.tmp)){
      if(bl.tmp$dir[irow] == '+'){
        orientation.target[bl.tmp$V8[irow]] = orientation.target[i]
      } else {
        orientation.target[bl.tmp$V8[irow]] = setdiff(dir.seq, orientation.target[i])
      }
    }
    if(sum(orientation.target == '.') == 0) break
  }
  table(orientation.target)
  # 
  for(i in which(orientation.target == '-')){
    seqs.target[i] = revComplSeq(seqs.target[i])
  }
  
  # seqs.target.msa = unlist(lapply(seqs.target, function(s) paste0(s, collapse = '')))
  # seqs.target.msa <- DNAStringSet(seqs.target.msa)
  # alignment <- msa(seqs.target.msa)
  
  # Run the alignment
  
  tmp.fasta = paste(path.mafft, 'tmp_',i.comp,'.fasta', sep = '')
  aln.fasta = paste(path.mafft, 'aln_',i.comp,'.fasta', sep = '')
  
  writeFasta(seqs.target, tmp.fasta)
  
  pokaz('Mafft...')
  system(paste('mafft --op 5 --quiet --maxiterate 100 ', tmp.fasta, '>', aln.fasta,  sep = ' '))
  
  pokaz('Plot...')
  alignment = readFasta(aln.fasta)
  
  seqs.mx = aln2mx(alignment)
  seqs.mx = seqs.mx[order(rowSums(seqs.mx == '-')),]
  p.msa = msaplot(seqs.mx) + annotate("text", x = -Inf, y = Inf, 
                                      label = i.comp, hjust = 0, vjust = 1)
  
  
  p.msadiff = msadiff(seqs.mx) + annotate("text", x = -Inf, y = Inf, 
                                      label = i.comp, hjust = 0, vjust = 1)
  
  
  # p.msa
  # p.msadiff
  
  pokaz('Save...')
  png(paste(path.figures, 'comp_', sprintf("%03d", i.comp),'_msa.png', sep = ''),
      width = 7, height = 8, units = "in", res = 300)
  grid.arrange(p.msa, p.msadiff, nrow = 2)  
  dev.off()
}
  
  
  
  
  # p.msa = p.msa + annotate("text", x = -Inf, y = Inf, 
  #                                          label = i.target, hjust = 0, vjust = 1)
  # 
  # rownames(sv.se) = sv.se$gr
  # f.sv = sv.se[sub("\\|.*$", "", names(alignment)),]$freq.max
  # 
  # 
  # pdf(paste(path.components, 'graph_mob_component_', sprintf("%03d", comp.target),'_msa.pdf', sep = ''), 
  #     width = 6, height = 4)
  # 


```





# Separate components
```{r}
# i.comp = 29
i.seq = which(memb.comp == i.comp)

names.comp = names(memb.comp)[i.seq]


s = seqs[names.comp[5]]
# dotself.s(s, 15, 12)
# 
# dotself.s(s, 15, 10)

dotplot.s(s, s, 15, 12)


# s = seqs["SVgr_1_id_03982|1742"]

```


```{r}

s1 = seqs[names.comp[1]]
s2 = seqs[names.comp[2]]

g.content$edges.small[g.content$edges.small[,1] %in% c(names.comp[1], names.comp[2]),]

dotplot.s(s1, s2, 15, 13)

```


```{r}
g.comp$membership[g.comp$membership == 1]

seqs.target = seqs[names(memb.comp)]

writeFasta(seqs.target, paste0(path.sv, 'seq_sv_big_target.fasta'))
```


# ORF
```{r}

orf.res = orfFinder(s)

orfplot(orf.res$pos)

head(orf.res$orf)

orf.res$orf[1]


```

#### comp 7 comp 5
```{r}

s1 = s


dotplot.s(s, s1, 15, 12)

```







